\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-04-25 paper/main.tex
.

We only use proteins in our analysis that have at least one link in STRING or one association in PhenomeNET, and drugs with at least one side effect. Therefore, the intersection between these
resources yields 1,428 drugs and 7,368 human proteins with 32,212
interactions for STITCH, 1,837 interactions between 680 drugs and 1,458 proteins for Yamanishi, and 6,498 links between 949 drugs and 2,221 proteins for BioSnap dataset. We provide links to and methods
for obtaining and processing the necessary data on Github.

@@

As the difference in performance with different splitting schemes is
quite large, we further evaluated additional drug--target interaction
and drug--target affinity prediction methods that were trained and
evaluation on other datasets.  Following the results of MolTrans
\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared it to our
``na\"ive'' predictor as well as \name{} (see Table
\ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
pair and the protein split; we were able to reproduce the MolTrans
results (Table \ref{tab:comparison2}), showing a substantial
difference based on the splitting scheme. We additionally computed the
$MicroAUC_p$ score for all considered methods, leading to similar 
results as observed on the Yaminishi dataset. As \name{} considers less drugs than alternative approaches, we execute a one-tailed T-test with $\alpha=0.01$ over all considered methods. Considering Macro AUCs, \name{} improves significantly over MolTrans ($p<0.0001$), DeepDTI ($p<0.0001$), DeepDTA ($p<0.0001$), and DeepConv-DTI ($p<0.0001$) with respect to BioSnap dataset, and significantly over DTINet ($p<0.0001$) and DTIGEMS+ ($p<0.0001$) with respect to Yamanishi dataset. However, DTI-CDF performs significantly better than \name{} ($\alpha=0.05, p=0.02$) when considering Macro AUC. As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF. We further run another one-tailed T-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction. We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$) regarding Yamanishi dataset.

>>>>

We only use proteins in our analysis that have at least one link in
STRING or one association in PhenomeNET, and drugs with at least one
side effect. Therefore, the intersection between these resources
yields 1,428 drugs and 7,368 human proteins with 32,212 interactions
for STITCH, 1,837 interactions between 680 drugs and 1,458 proteins
for Yamanishi, and 6,498 links between 949 drugs and 2,221 proteins
for BioSnap dataset. We provide links to and methods for obtaining and
processing the necessary data in our Github repository.

@@

As the difference in performance with different splitting schemes is
quite large, we further evaluated additional drug--target interaction
and drug--target affinity prediction methods that were trained and
evaluation on other datasets.  Following the results of MolTrans
\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared it to our
``na\"ive'' predictor as well as \name{} (see Table
\ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
pair and the protein split; we were able to reproduce the MolTrans
results (Table \ref{tab:comparison2}), showing a substantial
difference based on the splitting scheme. We additionally computed the
$MicroAUC_p$ score for all considered methods, leading to similar 
results as observed on the Yaminishi dataset. As \name{} considers
less drugs than alternative approaches, we execute a one-tailed T-test
with $\alpha=0.01$ over all considered methods. Considering Macro
AUCs, \name{} improves significantly ($p<0.0001$, one-sided t-test)
over MolTrans, DeepDTI, DeepDTA, and DeepConv-DTI on the BioSnap
dataset, and significantly ($p<0.0001$, one-sided t-test) over DTINet
and DTIGEMS+ on the Yamanishi dataset. However, DTI-CDF performs
significantly better than \name{} when considering Macro AUC ($p=0.02$). As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF. We further run another one-tailed T-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction. We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$) regarding Yamanishi dataset.
